---
title: "Week2 C5 assignment"
author: "letspairup"
date: "6/21/2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data and create data frame

- download data from source
- unzip it
- convert it to data frame


```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
              destfile = "activity.zip", mode="wb")

unzip("activity.zip")

repdata <- read.csv("activity.csv", header = TRUE)

```

## Sample data

```{r}
knitr::kable(head(repdata))
```

## Total number of steps by day

```{r, include=TRUE, message=FALSE}

library("dplyr")

dailystep_count <- repdata %>% 
  select(date, steps) %>% 
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarise(total_steps = sum(steps))

head(dailystep_count)

with(dailystep_count, 
     hist(total_steps, breaks = 15))

summary(dailystep_count)

```

High level summary of daily step counts. You can see __mean is 10766__ and __median is 10765__.

## Average daily activity pattern - Time series plot

```{r}
library(ggplot2)
repdata_by_interval <- repdata %>% na.omit() %>%
                      group_by(interval) %>% summarise(avg_steps= mean(steps)) 

ggplot(repdata_by_interval, aes(x=interval, y=avg_steps))+ geom_line()
```

## Five minute inteval with max steps - across all days

```{r}
repdata_by_interval[which.max(repdata_by_interval$avg_steps),]
```

## Total number of missing values

```{r}
sum(is.na(repdata))
```

## Total number of missing values

```{r}
sum(is.na(repdata))
```

## Strategy to fill missing values

Here we are going to use daily mean to fill missing values.

```{r}
fill_with_mean <- function(x)
{
  replace(x, is.na(x), mean(x, na.rm = TRUE))
}

complete_data <- repdata %>% mutate(steps = fill_with_mean(steps))

sum(is.na(complete_data))
```

complete_data is not having any missing values.:)

## Total number of steps by day and mean/median with complete data set

```{r}
dailystep_count_with_complete_data <- complete_data %>% 
  group_by(date) %>%
  summarise(total_steps = sum(steps))

with(dailystep_count_with_complete_data, 
     hist(total_steps, breaks = 15))

summary(dailystep_count_with_complete_data)
```

You can see __mean is 10766__ and __median is 10766__. *__Not much different from when we removed NA values.__*

## Activity pattern between weekdays and weekend

- Convert text based date column to Date
- Add new column day_category; marked row as either weekend or weekday based on the day of date.
- Group by data by day_category and interval
- Calculate average steps in avg_steps

```{r}
str(repdata)
repdata$date <- as.Date(repdata$date)
str(repdata)


get_day_category <- function(x) {
  if("Saturday" == weekdays(x) | "Sunday" == weekdays(x))
  {
    return("weekend")
  }
  else
  {
    return("weekday")
  }
}

comparedata <- repdata %>% filter(!is.na(steps)) %>%
  mutate(day_category = sapply(date,get_day_category)) %>%
  select(steps,interval, day_category) %>%
  group_by(day_category, interval) %>%
  mutate(avg_steps = mean(steps, na.rm = TRUE))

#str(comparedata)
```
Plot time series graph
```{r}
ggplot(comparedata, aes(x=interval, y=avg_steps, color=day_category)) + geom_line() +
facet_grid(day_category ~.) + xlab("Time interval") + ylab("Average steps - Mean") +
    ggtitle("Comparison of steps between weekend and weekdays")
```

We can clearly see that more steps are taken in week days on an average. Also weekend data is well distributed across active time intervals.



